---
title: "class8: Hands on with PCA"
author: "Han Duong"
date: "April 25, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#PCA of UK food data
```{r}
 x <- read.csv("UK_foods.csv", row.names=1)
head(x)

barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))

pairs(x, col=rainbow(10), pch=16)
#prcomp expects observations to be rows and variables to be columns
# t() used to transpose function
pca <- prcomp(t(x))
#indicates that 97% of the variance is accounted for by PCA1 and PCA2
summary(pca)

#generate a PC1 vs. PC2 plot
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2", xlim = c(-270, 500))
text(pca$x[,1], pca$x[,2], colnames(x), col = c("orange", "red", "blue", "green"))

```
#Calculate how much variation in the original data each PC accounts for

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2)*100 )
v

# or try
z <- summary(pca)
z$importance
```
The information can be summarized in a plot of variance (eigenvalues) with respect to the principal component number (eigenvector number), which is gene below.

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```
```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

```{r}
## Lets focus on PC2 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```


##Biplots

```{r}
##The inbuilt biplot() can be useful for small datasets
biplot(pca)
```

#PCA of RNA-seq data
```{r}
rna.data <- read.csv("expression.csv", row.names=1)
head(rna.data)
```
How many genes and samples are in this data set? 
```{r}
genes <- nrow(rna.data)
samples <- ncol(rna.data)
genes
samples
```

```{r}
## Again we have to take the transpose of our data 
pca <- prcomp(t(rna.data), scale=TRUE)
 
## Simple un ploished plot of pc1 and pc2
plot(pca$x[,1], pca$x[,2])
```

```{r}
## Variance captured per PC 
pca.var <- pca$sdev^2

## Precent variance is often more informative to look at 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```
Scree Plot 
```{r}
barplot(pca.var.per, main="Scree Plot", 
        xlab="Principal Component", ylab="Percent Variation")
```
Making it a little more attractive
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(rna.data)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

text(pca$x[,1], pca$x[,2], labels = colnames(rna.data), pos=c(rep(4,5), rep(2,5)))
```

```{r}
## Another way to color by sample type
## Extract the first 2 characters of the sample name
sample.type <- substr(colnames(rna.data),1,2)
sample.type
```


```{r}
## now use this as a factor input to color our plot
plot(pca$x[,1], pca$x[,2], col=as.factor(sample.type), pch=16)
```

```{r}
loading_scores <- pca$rotation[,1]

## Find the top 10 measurements (genes) that contribute
## most to PC1 in either direction (+ or -)
gene_scores <- abs(loading_scores) 
gene_score_ranked <- sort(gene_scores, decreasing=TRUE)

## show the names of the top 10 genes
top_10_genes <- names(gene_score_ranked[1:10])
top_10_genes 
```

```{r}
sessionInfo()
```

