---
title: "Class12"
author: "Han Duong"
date: "May 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")

hiv <- read.pdb(file.name)

hiv
```

>Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure? 


```{r}
prot <- atom.select(hiv, "protein", value=TRUE)
write.pdb(prot,file="1hsg.protein.pdb")
prot
```
```{r}
lig <- atom.select(hiv, "ligand", value=TRUE)
write.pdb(lig, file="1hsg.ligand.pdb")
lig
```

We are now using AutoDockTools for docking purposes. Add on hydrogen interactions and discover binding site. 

- utilize autodock tools and vina

Now, we can visualize the docks and load it into VMD. 

```{r}
library (bio3d)

res <- read.pdb("all.pdbqt", multi=TRUE)

write.pdb(res, "results.pdb")
```

Assess the results quantitatively by calculating the RMSD (root mean square distance) between each of the docking results and the known crystal structure using the bio3d package.

```{r}
#multi=TRUE, makes it read the file as a multi model file
res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)

```

## Normal Mode analysis for flexibility prediction

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```

