---
title: 'Class 6: R functions'
author: "Han Duong"
date: "April 18, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About
The is my **class 6** R markdown document with my *code* and notes for the day

```{r}
plot(1:10, typ = "l", col="blue")
```

##Reading in files practice
```{r}
read.table("test1.txt", header = TRUE, sep =",")
```


```{r}
read.csv("test1.txt")
```

```{r}
read.table("test2.txt", header =TRUE, sep = "$")
```
```{r}
read.table("test3.txt")

```

You can read urls. 
```{r}
read.table("https://bioboot.github.io/bimm143_S19/class-material/test2.txt", header = TRUE, sep = "$")
```
##Functions

Functions have name, arguments (input to function), and a body (where the work gets done). 

```{r}
add <- function(x, y=1){
  # the body!
  x+y
}
```

You've just created a function which you can now use.

```{r}
add(10)
```

You can argue with the default function of y and change it to 10. 
```{r}
add(10, y=10)
```
Each number of the vector will be added by 1 because y was equal to 1. 
```{r}
add( c(1,2,3))
```
Y was originally a number so the function wouldn't work because y is made a string. 

add(1, "b")

```{r}
x <- c( 1, 10, "string")
is.numeric(x) 
range (x, na.rm=TRUE)
```


##Consider copy and paste errors:

df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))

df$b <- (df$b - **min(df$a))** / (max(df$b) - min(df$b))

df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))

df$d <- (df$d - min(df$d)) / **(max(df$a)** - min(df$d))

Writing a function: 

- Makes the purpose of the code more clear

- Reduce mistakes from copy/paste

- Makes updating your code easier

- Reduce duplication and facilitate re-use


You need a “name”, “arguments” and “body”…
*na.rm used to omit "NA"*
```{r}
rescale <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
# Test on a small example where you know the answer
rescale(1:10)
```
How would you fix this 
```{r}
#How would you get your function to work here...
rescale( c(1,2,NA,3,10))
```
```{r}
x <- ( c(1,2,NA,3,10))
 rng <-range(x, na.rm = TRUE)
 rng
 (x - rng[1]) / (rng[2] - rng[1])

```

```{r}
rescale2 <- function(x, na.rm=TRUE) {
  
  if(!is.numeric(x)){
    stop("You only can input numbers")
  }
 rng <-range(x, na.rm = na.rm)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2(c(1,10))
```

```{r}
rescale2(c(1,2,NA,3,10))
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
  #each string prints when part of the function is executed
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 
 print("is it me you are looking for?")
#if loop, executes only when plot = TRUE
  if(plot) {
     plot(answer, typ="b", lwd=4)
    print("Please don't sing again")
 }
 print("I can see it in ...")
 
 #return statement usually the last statement because any functions following it will be ignored. 
 return(answer)
 # it's not required but recommended because it returns to control to the code that called it
}
```

```{r}
rescale3( c(1,2,NA,10))
```

Notice it shows only part of it
```{r}
rescale3( c(1,2,NA,10), plot =TRUE)
```


## Hands On Work 






























