---
title: "Class13"
author: "Han Duong"
date: "May 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sample genotype in the MXL 1000 Genome project for asthma

Here we focus on the Mexican Ancestry in Los Angeles, California (MXL) population. 

Main question: Does the GG genotype have an affect on prevalence of asthma of the MXL population? 

What proportion of the Mexican Ancestry in LA sample population(MXL) are homozygous fot the asthma associated SNP (G|G) rs8067378
```{r}
#Read CSV from ENSEMBLE
mxl <- read.csv("SampleMXL.csv")
head(mxl)

```

How many of each genotype are there?
```{r}
table(mxl$Genotype..forward.strand.)
```

Proportion or percent of total for each genotype
```{r}
(table(mxl$Genotype..forward.strand.) / nrow(mxl)) * 100
```

## Quality Scores in FASTQ fiels

The forth line of a FASTQ sequence format file encodes the quality score that tells us how food the sequence at a give position is (i.e. how likely it is to be correct based on the instrument)

```{r}
library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred
```

##Population Analytics
```{r}
pop <- read.table("popanalytics.txt")
summary(pop)

```

```{r}
inds <- pop$geno == "G/G"
summary(pop[inds, "exp"])
```

```{r}
inds <- pop$geno == "A/G"
summary(pop[inds, "exp"])
```

```{r}
inds <- pop$geno == "A/A"
summary(pop[inds, "exp"])
```

```{r}
boxplot(exp ~ geno ,data=pop)
```

