---
title: 'Class7: R functions and packages'
author: "Han Duong"
date: "April 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions revisted

We will source a file from online with our function from last slide. 

```{r}
source("http://tinyurl.com/rescale-R")

```

Try out the last day's rescale() function 

```{r}
rescale(1:10)
```

warning() and stop() functions are used inside functions to handle and report on unexpected situations

"Fail early and loudly"


Try rescale2() function that catches string inputs

```{r}
rescale2(1:10, "string")
```

#Find missing NA values in two vectors
Now we're going to write a function both_na(), that counts how many positions in two input vectors, x and y, both have a missing value
1. Always with a simple definition of the problem

```{r}
# Lets define an example x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

#Use the is.na() function (can be found by simply searching "r find missing values in a vector")
is.na(x)
is.na(y)
#this function reports TRUE when it encounters NA in a vector, but how do you get it to report only NA present in both vectors
```

Try putting these together with an AND
```{r}
is.na(x) & is.na(y)
```

Take the sum() to find out how many TRUE NAs we have and thus how many NAs we had in both x and y

```{r}
#our working snippet
sum(is.na(x) & is.na(y))
```

Now I can make this into our first function

```{r}
both_na <- function(x, y){
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test, test, test
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

This function recycles the first value of x to make it the same length as y2. Which gives  x <- c(NA, NA, NA, NA), resulting in "3". Different vector lengths might result in an inaccurate value. 
```{r}
both_na(x, y2)
```

```{r}
3==3
```


Now let's try the bothna2() function on our different length input vectors

```{r}
both_na2(x, y2)
```

```{r}
which(c(F, F, T, F, T))


#which (is.na(c(1, 2, NA, 4)))
```

## intersect function

```{r}
df2
```
Make things simple, turn them into single vectors

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

R intersection of two vectors

```{r}
intersect(x, y)
```

```{r}
which (x %in% y)
```

```{r}
x[x %in% y]
```

Use the RStudio shortcut 'CODE > EXTRACT FUNCTION' to turn our snippet into a working function (highlighted)

```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[ y%in% x])
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```


```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

# Create a function 

grade() that finds average grade dropping the worst homework score
```{r}
x <- c(100,100,100,100,100,100,100,90)
y <- c(100,90,90,90,90,90,97,80)

grade <- function(x) {
  score <- (sum(x) - min(x))/ (length(x)-1)
if (score < 90){
  paste("You got",score,".", " Work harder!")
}
}
```


```{r}
grade(c(70,50,40,80,90,100))
```






























